---
title: "BASIX project"
output: slidy_presentation
date: "2022-12-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Summer 2003, policy marketed to 2 villages using BASIXs best sales force 148 farmers purchased the policy# Paying 74000 Rs (1800\$) for coverage of 600 total acres of land 15% commission on sales, negligible revenue far outweighed by staff time, developing customer relations

BASIX at large 60000 borrowers in 4000 villages in 30 districts

loan products typically had an annual percentage rate of 24% primarily individual, sometimes grouped joint liabilities. average load size of 8800 Rs,

typically repaid in 11 to 24 months, monthly repayments matched to cash flow? crop loans mature after harvest small enterprise loans require regular month payments

BASIX primary source of operating income is interest on the above loans SEE INCOME EXHIBIT 2 9% cost of capitol 2% provisioning for future loan losses FY 2003 expected long term loan loss rate was \~ 2% typical repayment rates for micro finance institutions ranged between 97% and 99% BASIX has a lower repayment rate then typical. Because of drought

BASIX also sells life insurance, first to borrowers, bundled with the loan, 60000 borrowers policy covered 1.5 times the amount of the initial loan and eventually even to non-borrowers By 2004 BASIX is using 250 customer service agents serving exclusively areas where commercial banks and micro finance corporations were not operating areas previously served by local moneylenders 10% interest per month or higher no competition for insurance sales for at least 5-10 years

A BASIX client agriculture as a primary source of income in Andhra Pradesh (AP) a single household of \~5 members, a couple, a parent, and two children. \~3 acres of land ownership whereupon they would grow groundnuts or castor seed. 3 acres of groundnut yielded 1 ton (1000 kg) of seeds, which could be sold for 25-35 Rs / kg. middle men take substantial cut of profits households also have farm animals, grow food crops (maize or sorghum) expenses include farming inputs, seed and fertilizer, food, clothing, and fuel

SEE CLIENTELE INCOME DISTRIBUTION IN EXHIBIT 3

Farmers in AP and Tamil Nadu grow rice, groundnuts, sorghum, and more? limited to no irrigation rainfall levels are an important determinant of agricultural productivity The quality of monsoons is an actual determinant of economic growth nationwide two growing cycles Jun- Nov and Dec- Apr monsoon occurs during Jun- Nov

Farmers had to pay for seeds, fertilizer and labor long before profits. Goods were purchased on credit so farmers are in debt most of the year.

HOUSE HOLD RISK MANAGEMENT Grim situation with perilous, costly circumstances, harsh saving conditions.

FINANCIAL RISK MANAGEMENT SERVICES Helping the house hold risk management situation needed a credit solution. However a credit solution was not enough. Provisions for insurance against livestock death and crop failure would go far.

Crop insurance classically sold by the national government was a conditioned bundling for borrowing. BASIX clients are ineligible.

National crop insurance was no premier service though. Unequally distributed benefits and slow payments.

Earliest crop insurance product by BASIX in 2000, 3 villages, farmers pooled risk, individuals would pay a premium annually. A fraction of the funds would be kept in a village account. Balance accumulated in a "multi-village" account. Farmers would internally accept or reject claims. Losses would be drawn from the village account, and then the "mutli-village" account. BASIX was liable for claims exceeding the total premiums accumulated.

Administration costs were too high. Farmers were reluctant to pay the high premiums.

RAINFALL INSURANCE

Using a designated rainfall station near clients: Policy would pay cash if measured rainfall was particularly low. Clients would pay premiums.

A private insurance company was brought in with international connections. BASIX would act only as a sales agent, and earn commissions. Not be responsible for the payouts.

**How should the rainfall index be designed?**

**What should the targets and payouts be?**

**What would be the policy's price?**

**Specific policies per crop?**

**One-size-fits all?**

**Which would be easier to sell?**

PRODUCT DESIGN AND MARKETING

Pilot program in only a few of the villages. Mahbudnagar district of AP.

2003 Pilot policy would target just groundnuts and castor seed.

Food and Agricultural Organization suggested that groundnut required approximately 500-700 millimeters of water per year to produce good yield.

Growing period of May/June - mid October.

**A MODEL FOR THE RELATIONSHIP BETWEEN CROP LOSS AND RAINFALL DEFICIT**

**blah blah blah**

Weights are assigned to each part of the growing season to calculate rainfall index.

Insurance policy sets 653 mm target rainfall *index* over the growing period: May 11 to October 17, 2003.

**SEE EXHIBIT 5 for summary of 2003 policy**

BASIX ignores other crop categories and focuses only on groundnuts and castor seed.

The policy only sold 150 policies.

Fixed costs of development and administration and the cost of field agents' time far outweighed first year profits.

Agents took 3 hours to sell 600 Rs insurance policies,

Agents took 6 hours to sell make and service an average size loan.

Neglecting customer service agent travel time.

THE 2004 REVISED PRODUCT

Farmers were concerned about transparency.

**SEE EXHIBIT 5 for revised policy**

**SEE EXHIBIT 6 for historical distribution of rainfall for the periods covered by the policy.**

DECISION

2003 rainfall resulted in modest payouts.

Farmers were satisfied about the speed of payouts.

Good interest in policy for next monsoon season.

Caution was important though, trust was long earned.

Bundling the insurance and the loans could result in misunderstanding and default.

**Would the policy sell well?**

**What was the best way to explain it to clients?**

**How to avoid "gouging' its captive clients?**

**Was the policy fair?**

**How could it help BASIX clients?**

**WHAT WERE THE OTHER IMPORTANT RISKS, COSTS, or BENEFITS that hadn't been considered?**

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

-   Bullet 1
-   Bullet 2
-   Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```
